Statically Typed String Sanitation Inside a Python (Technical Report)

### 2014-12-16 sanitation-psp14/tr/strings_chapter.tex
fixed some errors.

Our proofs here are more sketchy than in the $\lambdas$ section both because the
proof method is already demonstrated, and becaues $\lambdap$ is a fairly straightforward
extension of $\lambda$ for which these results are already estbalished.

@@

\begin{thm}[Progress]
If $\emptyset\vdash \iota : \tau$ either $\iota = \dot{v}$ or $\iota \mapsto \iota'$ for some $\iota'$.
\end{thm}
\begin{proof}
The proof proceeds by induction on the typing assumption. We elide contexts where context has obviously not changed.

Consider only the string and regex (non-$\lambda$) fragments of $\lambdap$.

Regular expressions are always values.

For strings, the concatenation cases are trivial.

@@

\item \textbf{Case SP-E-ConcatLeft}.
Suppose:
\begin{align*}
  \iota &= \concat{\iota_1}{\iota_2} \mapsto \concat{\iota_1'}{\iota_2} \\
  \emptyset &\vdash \iota : \str \\
  \iota &\mapsto \iota' 
\end{align*}

The only appliable typing rule is P-T-Concat, so $\emptyset \vdash \iota_1 : \str$
and $\emptyset \vdash \iota_2 : \str$ where $\tau = \str$. 
By induction, $\emptyset \vdash \iota_1' : \str$,
so $\emptyset \vdash \concat{\iota_1'}{\iota_2} : \str$.

\item \textbf{Case SP-E-ConcatRight}
\begin{align*}
  e &= \concat{e_1}{e_2} \mapsto \concat{e_1}{e_2'} \\
  \emptyset &\vdash e : \str \\
  \iota &\mapsto \iota' 
\end{align*}

The only appliable typing rule is P-T-Concat, so $\emptyset \vdash \iota_1 : \str$
and $\emptyset \vdash \iota_2 : \str$ where $\tau = \str$. 
By induction, $\emptyset \vdash \iota_1' : \str$,
so $\emptyset \vdash \concat{\iota_1'}{\iota_2} : \str$.

@@

\begin{mathpar}
\inferrule[PP-E-ConcatLeft]
{\iota_1 \mapsto \iota_1'}
{\tconcat{\iota_1}{\iota_2} \treduces \tconcat{\iota_1'}{\iota_2}}

\inferrule[PP-E-ConcatRight]
{\iota_2 \mapsto \iota_2'}
{\tconcat{\iota_1}{\iota_2} \treduces \tconcat{\iota_1}{\iota_2'}}

\inferrule[PP-E-Concat]
{\ }
{\tconcat{\tstr{s_1}}{\tstr{s_2}} \treduces \tstr{s_1s_2}}

\inferrule[PP-E-CaseLeft]
{ \iota_1 \mapsto \iota_1' }
{ \pstrcase{\iota_1}{\iota_2}{x,y.\iota_3} \mapsto \pstrcase{\iota_1'}{\iota_2}{x,y.\iota_3} }

\inferrule[PP-E-CaseRight]
{ \iota_2 \mapsto \iota_2' }
{ \pstrcase{\dot{v}}{\iota_2}{x,y.\iota_3} \mapsto \pstrcase{\dot{v}}{\iota_2'}{x,y.\iota_3} }

\inferrule[PP-E-Case]
{ }
{ \pstrcase{\tstr{s_1}}{\tstr{s_2}}{x,y.\iota_3} \mapsto \tstr{s_1s_2}}

\inferrule[PP-E-ReplaceLeft]
{\iota_1 \mapsto \iota_1'}
{ \preplace{\iota_1}{\iota_2}{\iota_3} \mapsto \preplace{\iota_1'}{\iota_2}{\iota_3} }

\inferrule[PP-E-ReplaceMid]
{\iota_2 \mapsto \iota_2'}
{ \preplace{\rx{r}}{\iota_2}{\iota_3} \mapsto \preplace{\rx{r}}{\iota_2'}{\iota_3} }

\inferrule[PP-E-ReplaceRight]
{\iota_3 \mapsto \iota_3'}
{ \preplace{\rx{r}}{\tstr{s_2}}{\iota_3} \mapsto \preplace{\rx{r}}{\tstr{s_2}}{\iota_3'} }

\inferrule[PP-E-Replace]
{\ }
{ \preplace{\rx{r}}{\tstr{s_2}}{\tstr{s_3}} \mapsto \tstr{\lsubst{r}{s_2}{s_3}} }

\inferrule[PP-E-Check-Left]
{ \iota_x \mapsto \iota_x' }
{ \check{\iota_x}{\iota}{\iota_1}{\iota_2} \mapsto \check{\iota_x}{\iota}{\iota_1}{\iota_2} }

\inferrule[PP-E-Check-Right]
{ \iota \mapsto \iota' }
{ \check{\rx{r}}{\iota}{\iota_1}{\iota_2} \mapsto \check{\rx{r}}{\iota'}{\iota_1}{\iota_2} }

\inferrule[PP-E-Check-OK]
{ s \in \lang{r} }
{ \check{\rx{r}}{\tstr{s}}{\iota_1}{\iota_2} \mapsto \iota_1}

\inferrule[PP-E-Check-OK]
{ s \not\in \lang{r} }
{ \check{\rx{r}}{\tstr{s}}{\iota_1}{\iota_2} \mapsto \iota_2}

>>>>

@@

\begin{thm}[Progress]
If $\emptyset\vdash \iota : \tau$ either $\iota = \dot{v}$ or $\iota \mapsto \iota'$ for some $\iota'$.
\end{thm}
\begin{proof}
The proof proceeds by induction on the typing assumption. 
Consider only the string and regex (non-$\lambda$) fragments of $\lambdap$.
\begin{itemize}[label=$ $,itemsep=1ex]

@@

\item \textbf{Case PS-E-ConcatLeft}.
Suppose:
\begin{align*}
  \iota &= \rconcat{\iota_1}{\iota_2} \mapsto \rconcat{\iota_1'}{\iota_2} \\
  \emptyset &\vdash \iota : \str \\
  \iota &\mapsto \iota' 
\end{align*}

The only applicable typing rule is P-T-Concat, so $\emptyset \vdash \iota_1 : \str$
and $\emptyset \vdash \iota_2 : \str$ where $\tau = \str$. 
By induction, $\emptyset \vdash \iota_1' : \str$,
so $\emptyset \vdash \rconcat{\iota_1'}{\iota_2} : \str$.

\item \textbf{Case PS-E-ConcatRight}
\begin{align*}
  e &= \rconcat{e_1}{e_2} \mapsto \rconcat{e_1}{e_2'} \\
  \emptyset &\vdash e : \str \\
  \iota &\mapsto \iota' 
\end{align*}

The only applicable typing rule is P-T-Concat, so $\emptyset \vdash \iota_1 : \str$
and $\emptyset \vdash \iota_2 : \str$ where $\tau = \str$. 
By induction, $\emptyset \vdash \iota_1' : \str$,
so $\emptyset \vdash \rconcat{\iota_1}{\iota_2'} : \str$.

\todo{Everything above in this proof was sleep deprived plane work -- need to check}

\item \textbf{Case PS-E-Concat}
Let $e = \rconcat{\rstr{s_1}}{\rstr{s_2}} \mapsto \rstr{s_1s_2}$. 
The only rule that applies is P-T-Concat, so $\theta \vdash e : \str$.
By canonical forms, $\theta \rstr{s_1s_2} : \str$.

\item \textbf{Case PS-E-CaseLeft}
Let $\iota = \strcase{\iota_1}{\iota_2}{x,y.\iota_3} \mapsto \strcase{\iota_1'}{\iota_2}{x,y.\iota_3}$
when $\iota_1 \mapsto \iota_1'$.
The only rule that applies is P-T-Case, so $\theta \vdash \iota : \tau$ where:
\begin{align*}
  \theta &\vdash \iota_1 : \str 
  \theta &\vdash \iota_2 : \tau
  \theta, x : \str, y : \str &\vdash \iota_3 : \tau
\end{align*}
By induction, $\theta \vdash \iota_1' : \str$.
By P-T-Case, $\thet \vdash \strcase{\iota_1}{\iota_2}{x,y.\iota_3} : \tau$.

\item \textbf{Case PS-E-CaseRight}
Let $\iota = \strcase{\iota_1}{\iota_2}{x,y.\iota_3} \mapsto \strcase{\iota_1}{\iota_2}{x,y.\iota_3}$
when $\iota_1 \mapsto \iota_1'$.
The only rule that applies is P-T-Case, so $\theta \vdash \iota : \tau$ where:
\begin{align*}
  \theta &\vdash \iota_1 : \str 
  \theta &\vdash \iota_2 : \tau
  \theta, x : \str, y : \str &\vdash \iota_3 : \tau
\end{align*}
By induction, $\theta \vdash \iota_1' : \str$.
By P-T-Case, $\thet \vdash \strcase{\iota_1}{\iota_2}{x,y.\iota_3} : \tau$.

@@

\begin{figure}[t]
\small
$\fbox{\inferrule{}{\iota \mapsto \iota}}$

\begin{mathpar}
\inferrule[PS-E-ConcatLeft]
{\iota_1 \mapsto \iota_1'}
{\tconcat{\iota_1}{\iota_2} \treduces \tconcat{\iota_1'}{\iota_2}}

\inferrule[PS-E-ConcatRight]
{\iota_2 \mapsto \iota_2'}
{\tconcat{\iota_1}{\iota_2} \treduces \tconcat{\iota_1}{\iota_2'}}

\inferrule[PS-E-Concat]
{\ }
{\tconcat{\tstr{s_1}}{\tstr{s_2}} \treduces \tstr{s_1s_2}}

\inferrule[PS-E-CaseLeft]
{ \iota_1 \mapsto \iota_1' }
{ \pstrcase{\iota_1}{\iota_2}{x,y.\iota_3} \mapsto \pstrcase{\iota_1'}{\iota_2}{x,y.\iota_3} }

\inferrule[PS-E-Case-Epsilon]
{  }
{ \pstrcase{\epsilon}{\iota_2}{x,y.\iota_3} \mapsto \iota_2}

\inferrule[PS-E-Case]
{ }
{ \pstrcase{\tstr{as}}{\iota_2}{x,y.\iota_3} \mapsto \tstr{as}}

\inferrule[PS-E-ReplaceLeft]
{\iota_1 \mapsto \iota_1'}
{ \preplace{\iota_1}{\iota_2}{\iota_3} \mapsto \preplace{\iota_1'}{\iota_2}{\iota_3} }

\inferrule[PS-E-ReplaceMid]
{\iota_2 \mapsto \iota_2'}
{ \preplace{\rx{r}}{\iota_2}{\iota_3} \mapsto \preplace{\rx{r}}{\iota_2'}{\iota_3} }

\inferrule[PS-E-ReplaceRight]
{\iota_3 \mapsto \iota_3'}
{ \preplace{\rx{r}}{\tstr{s_2}}{\iota_3} \mapsto \preplace{\rx{r}}{\tstr{s_2}}{\iota_3'} }

\inferrule[PS-E-Replace]
{\ }
{ \preplace{\rx{r}}{\tstr{s_2}}{\tstr{s_3}} \mapsto \tstr{\lsubst{r}{s_2}{s_3}} }

\inferrule[PS-E-Check-Left]
{ \iota_x \mapsto \iota_x' }
{ \check{\iota_x}{\iota}{\iota_1}{\iota_2} \mapsto \check{\iota_x}{\iota}{\iota_1}{\iota_2} }

\inferrule[PS-E-Check-Right]
{ \iota \mapsto \iota' }
{ \check{\rx{r}}{\iota}{\iota_1}{\iota_2} \mapsto \check{\rx{r}}{\iota'}{\iota_1}{\iota_2} }

\inferrule[PS-E-Check-OK]
{ s \in \lang{r} }
{ \check{\rx{r}}{\tstr{s}}{\iota_1}{\iota_2} \mapsto \iota_1}

\inferrule[PS-E-Check-NotOK]
{ s \not\in \lang{r} }
{ \check{\rx{r}}{\tstr{s}}{\iota_1}{\iota_2} \mapsto \iota_2}
